\documentclass{ximera}

\title{Practice for Separable ODEs}

\input{../../preamble}

\begin{document}
\begin{abstract}
    Why?
\end{abstract}
\maketitle


\begin{exercise}
    Solve $y' = y^3$ for $y(0)=1$.
\end{exercise}
%\comboSol{%
%}
%{%
%$y = \frac{1}{\sqrt{1-2x}}$
%}

\begin{exercise}%
    Solve $x' = \frac{1}{x^2}$, $x(1)=1$.
\end{exercise}
%\exsol{%
%$x = {(3t-2)}^{1/3}$
%}


\begin{exercise}[little harder]
    Solve $y' = (y-1)(y+1)$ for $y(0)=3$. (Note: Requires partial fractions)
\end{exercise}
%\comboSol{%
%}
%{%
%$y = \frac{2 + e^{2t}}{2 - e^{2t}}$
%}

\begin{exercise}%
    Solve $x' = \frac{1}{\cos(x)}$, $x(0)=\frac{\pi}{2}$.
\end{exercise}
%\exsol{%
%$x = \sin^{-1} \bigl(t+1\bigr)$
%}


\begin{exercise}
    Solve $\frac{dy}{dx} = \frac{1}{y+1}$ for $y(0)=0$.
\end{exercise}
%\comboSol{%
%}
%{%
%$y = -1 + \sqrt{1+2x}$
%}

\begin{exercise}
    Solve $y' = \nicefrac{x}{y}$.
\end{exercise}
%\comboSol{%
%}
%{%
%$y = \pm \sqrt{2C + x^2}$
%}

\begin{exercise}
    Solve $y' = x^2y$.
\end{exercise}
%\comboSol{%
%}
%{%
%$y = De^{\frac{x^3}{3}}$
%}

\begin{exercise}%
    Consider the differential equation \[ \frac{dy}{dx} = \frac{2x}{y} \]
    \begin{tasks}
        \task Find the general solution as an implicit function.
        \task Find the solution to this differential equation as an \emph{explicit} function with $y(1) = 4$.
        \task Find the solution to this differential equation as an \emph{explicit} function with $y(0) = -2$. 
    \end{tasks}
\end{exercise}
%\exsol{%
%a) $\frac{y^2}{2} = x^2 + C$ \quad b) $y = \sqrt{2x^2 + 14}$ \quad c) $y = -\sqrt{2x^2 + 4}$
%}

\begin{exercise}
    Solve $y' = y^n$, $y(0) = 1$, where $n$ is a positive integer. Hint: You have to consider different cases.
\end{exercise}
%\exsol{%
%If $n \not= 1$, then
%$y={\bigl((1-n)x+1\bigr)}^{1/(1-n)}$.
%If $n=1$, then $y = e^x$.
%}


\begin{exercise}
    Solve $\dfrac{dx}{dt} = (x^2-1)\,t$, for $x(0) = 0$. (Note: Requires partial fractions)
\end{exercise}
%\comboSol{%
%}
%{%
%$x = \frac{1 - e^{t^2}}{1 + e^{t^2}}$
%}

\begin{exercise}
    Solve $\dfrac{dx}{dt} = x\,\sin(t)$, for $x(0) = 1$.
\end{exercise}
%\comboSol{%
%}
%{%
%$x = e^{1-\cos(t)}$
%}

\begin{exercise}%
    Solve $y'=2xy$.
\end{exercise}
%\exsol{%
%$y = Ce^{x^2}$
%}

\begin{exercise}
    Solve $y' = ye^{2x}$ with $y(0) = 4$. 
\end{exercise}
%\comboSol{%
%}
%{%
%$y = 4e^{\frac{1}{2}(e^{2x} - 1)}$
%}

\begin{exercise}
    Solve $\dfrac{dy}{dx} = xy+x+y+1$.  Hint: Factor the right-hand side.
\end{exercise}
%\comboSol{%
%}
%{%
%$y = Ce^{\frac{x^2}{2} + x} - 1$
%}

\begin{exercise}%
    Solve $x'=3xt^2-3t^2$, $x(0)=2$.
\end{exercise}
%\exsol{%
%$x = e^{t^3}+1$
%}

\begin{exercise}
    Find the general solution of $y' = e^x$,  and then $y' = e^y$.
\end{exercise}
%\comboSol{%
%}
%{%
%$y = e^x + C$, $y= = -\ln(|-x-C|)$
%}

\begin{exercise}
    Solve $xy' = y + 2x^2 y$, where $y(1) = 1$.
\end{exercise}
%\comboSol{%
%}
%{%
%$y = xe^{x^2 - 1}$
%}

\begin{exercise}%
    Find an implicit solution for $x'=\frac{1}{3x^2+1}$, $x(0)=1$.
\end{exercise}
%\exsol{%
%$x^3+x=t+2$
%}

\begin{exercise}
    Solve $\dfrac{dy}{dx} = \dfrac{y^2+1}{x^2+1}$, for $y(0) = 1$.
\end{exercise}
%\comboSol{%
%}
%{%
%$y = \tan\left(\tan^{-1}(x) + \frac{\pi}{4}\right)$
%}

\begin{exercise}
    Find an implicit solution for $\dfrac{dy}{dx} = \dfrac{x^2+1}{y^2+1}$, for $y(0) = 1$.
\end{exercise}
%\comboSol{%
%}
%{%
%$\frac{y^3}{3} + y = \frac{x^3}{3} + x + \frac{4}{3}$
%}

\begin{exercise}%
    Find an implicit solution to $y' = \frac{\sin(x)}{\cos(y)}$.
\end{exercise}
%\exsol{%
%$\sin(y) = -\cos(x) + C$
%}

\begin{exercise}
    Find an implicit solution for $xy' = \frac{x^2 + 1}{y^2 - 1}$ with $y(3) = 2$.
\end{exercise}
%\comboSol{%
%}
%{%
%$\frac{y^3}{3} - y = \frac{x^2}{2} +\ln(|x|) + 4 - \ln(2)$
%}

\begin{exercise}
    Find an explicit solution for $y' = xe^{-y}$, $y(0)=1$.
\end{exercise}
%\comboSol{%
%}
%{%
%$y = \ln\left(\frac{x^2}{2} + e \right)$
%}

\begin{exercise}%
    Find an explicit solution to $x y' = y^2$, $y(1) = 1$.
\end{exercise}
%\exsol{%
%$y = \frac{1}{1-\ln x}$
%}

\begin{exercise}
    Find an explicit solution for $xy' = e^{-y}$, for $y(1)=1$.
\end{exercise}
%\comboSol{%
%}
%{%
%$y = \ln|\ln(|x|) + e|$
%}

\begin{exercise}
    Find an explicit solution for $y' = y^2(x^4 + 1)$ with $y(1) = 2$. 
\end{exercise}
%\comboSol{%
%}
%{%
%$y = \frac{1}{\frac{17}{10} - x - \frac{x^5}{5}}$
%}

\begin{exercise}
    Find an explicit solution for $y' = \frac{\cos(x)+ 1}{y}$ with $y(0) = 4$. 
\end{exercise}
%\comboSol{%
%}
%{%
%$y = \sqrt{2\sin(x) + 2x + 16}$
%}

\begin{exercise}
    Find an explicit solution for $y' = ye^{-x^2}$, $y(0)=1$.  It is alright to leave a definite integral in your answer.
\end{exercise}
%\comboSol{%
%}
%{%
%$y = e^{\int_0^x e^{-s^2}\ ds}$
%}

\begin{exercise}
    Is the equation $y' = x+y+1$ separable? If so, find the general solution, if not, explain why.
\end{exercise}
%\comboSol{%
%}
%{%
%No
%}

\begin{exercise}
    Is the equation $y' = ty^2 + t$ separable? If so, find the general solution, if not, explain why.
\end{exercise}
%\comboSol{%
%}
%{%
%Yes. $y = \tan\left(\frac{t^2}{2} + C\right)$
%}

\begin{exercise}
    Is the equation $y' = xy^2 + 3y^2 - 4x - 12$ separable? If so, find the general solution, if not, explain why. (Note: Requires partial fractions)
\end{exercise}
%\comboSol{%
%}
%{%
%Yes. $\frac{1}{2\sqrt{3}}\ln\left(\left| \frac{y - \sqrt{3}}{y+\sqrt{3}}\right|\right) = \frac{x^2}{2} + 3x + C$
%}

\begin{exercise}
    Suppose a cup of coffee is at 100 degrees Celsius at time $t=0$, it is at 70 degrees at $t=10$ minutes, and it is at 50 degrees at $t=20$ minutes.  Compute the ambient temperature.
\end{exercise}
%\comboSol{%
%}
%{%
%10 C
%}

\begin{exercise}%
    Take \exampleref{sep:coffeeexample} with the same numbers: 89 degrees at $t=0$, 85 degrees at $t=1$, and ambient temperature of 22 degrees.  Suppose these temperatures were measured with precision of $\pm 0.5$ degrees.  Given this imprecision, the time it takes the coffee to cool to (exactly) 60 degrees is also only known in a certain range.  Find this range.  Hint: Think about what kind of error makes the cooling time longer and what shorter.
\end{exercise}
%\exsol{%
%The range is approximately 7.45 to 12.15 minutes.
%}

\begin{exercise}%
    A population $x$ of rabbits on an island is modeled by $x' = x- \bigl(\nicefrac{1}{1000} \bigr) x^2$, where the independent variable is time in months.  At time $t=0$, there are 40 rabbits on the island.
    \begin{tasks}
        \task Find the solution to the equation with the initial condition.
        \task How many rabbits are on the island in 1 month, 5 months,  10 months, 15 months (round to the nearest integer).
    \end{tasks}
\end{exercise}
%\exsol{%
%a) $x = \frac{1000e^t}{e^t+24}$. \quad b) 102 rabbits after one month,
%861 after 5 months, 999 after 10 months, 1000 after 15 months.
%}

%\setcounter{exercise}{100}


\end{document}