\documentclass{ximera}

\title{Practice for Second Order Linear ODEs}

\input{../../preamble}

\begin{document}
\begin{abstract}
    Why?
\end{abstract}
\maketitle


\begin{exercise}
    Show that $y=e^x$ and $y=e^{2x}$ are linearly independent.
\end{exercise}
%\comboSol{%
%}
%{%
%Ratio is non-constant.
%}

\begin{exercise}%
    Are $\sin(x)$ and $e^x$ linearly independent?  Justify.
\end{exercise}
%\exsol{%
%Yes.  To justify try to find a constant $A$ such that $\sin(x) = A e^x$
%for all $x$.
%}


\begin{exercise}%
    Are $e^x$ and $e^{x+2}$ linearly independent?  Justify.
\end{exercise}
%\exsol{%
%No.  $e^{x+2} = e^2 e^x$.
%}

\begin{exercise}%
    Guess a solution to $y'' + y' + y= 5$.
\end{exercise}
%\exsol{%
%$y=5$
%}

\begin{exercise}
    Take $y'' + 5 y = 10 x + 5$.  Find (guess!) a solution.
\end{exercise}
%\comboSol{%
%}
%{%
%$y = 2x+1$
%}

\begin{exercise}
    Verify that $y_1(t) = e^t \cos(2t)$ and $y_2(t) = e^t \sin(2t)$ both solve $y'' - 2y' + 5y = 0$. Are these two solutions linearly independent? What does that mean about the general solution to $y'' - 2y' + 5y = 0$?
\end{exercise}
%\comboSol{%
%}
%{%
%Yes. This means the general solution is $C_1e^{t}\cos(2t) + C_2\sin(2t)$.
%}

\begin{exercise}
    Prove the superposition principle for nonhomogeneous equations.  Suppose that $y_1$ is a solution to $L y_1 = f(x)$ and $y_2$ is a solution to $L y_2 = g(x)$ (same linear operator $L$).  Show that $y = y_1+y_2$ solves $Ly = f(x) + g(x)$.
\end{exercise}
%\comboSol{%
%}
%{%
%Hint: Plug in $y_1 + y_2$ and see that it works correctly.
%}

\begin{exercise}
    Determine the maximal interval of existence of the solution to the differential equation
    \[ 
        (t - 5)y'' + \frac{1}{t+1}y' + e^t y = \frac{\cos(t)}{t^2 + 1} 
    \] 
    with initial condition $y(3) = 8$. What about if the initial condition is $y(-3) = 4$?
\end{exercise}
%\comboSol{%
%}
%{%
%$(-1, 5)$ for $y(3) = 8$. $(-\infty, -1)$ for $y(-3) = 4$.
%}

\begin{exercise}
    For the equation $x^2 y'' - x y' = 0$, find two solutions, show that they are linearly independent and find the general solution.
Hint: Try $y = x^r$.
\end{exercise}
%\comboSol{%
%}
%{%
%$y = C_1 + C_2x^2$
%}

\begin{exercise}%
    Find the general solution to $x y'' + y' = 0$.  Hint: It is a first order ODE in $y'$.
\end{exercise}
%\exsol{%
%$y=C_1 \ln(x) + C_2$
%}

\begin{exercise}
    Find the general solution of $2y'' + 2y' -4 y = 0$.
\end{exercise}
%\comboSol{%
%}
%{%
%$y = C_1e^{-2t} + C_2e^{t}$
%}

\begin{exercise}
    Solve $y'' + 9y' = 0$ with $y(0) = 1$, $y'(0) = 1$.
\end{exercise}
%\comboSol{%
%}
%{%
%$y = \frac{10}{9} - \frac{1}{9}e^{-9t}$
%}
\begin{exercise}
    Find the general solution of $y'' + 9y' - 10 y = 0$.
\end{exercise}
%\comboSol{%
%}
%{%
%$y = C_1e^{-10t} + C_2e^t$
%}

\begin{exercise}
    Find the general solution to $y'' - 3y' - 4y = 0$.
\end{exercise}
%\comboSol{%
%}
%{%
%$y = C_1e^{4t} + C_2e^{-t}$
%}

\begin{exercise}
    Find the general solution to $y'' + 6y' + 8y = 0$.
\end{exercise}
%\comboSol{%
%}
%{%
%$y = C_1e^{-2t} + C_2e^{-4t}$
%}

\begin{exercise}
    Find the solution to $y'' - 3y' + 2y = 0$ with $y(0) = 3$ and $y'(0) = -1$.
\end{exercise}
%\comboSol{%
%}
%{%
%$y = 7e^t - 4e^{2t}$
%}

\begin{exercise}
    Find the solution to $y'' + y' -12y = 0$ with $y(0) = 1$ and $y'(0) = -2$.
\end{exercise}
%\comboSol{%
%}
%{%
%$y = \frac{5}{7}e^{-4t} + \frac{2}{7}e^{3t}$
%}
\begin{exercise}%
    Find the general solution to $y''+4y'+2y=0$.
\end{exercise}
%\exsol{%
%$y =
%C_1 e^{(-2+\sqrt{2}) x}
%+
%C_2 e^{(-2-\sqrt{2}) x}$
%}

\begin{exercise}%
    Find the solution to $2y''+y'-3y=0$, $y(0) = a$, $y'(0)=b$.
\end{exercise}
%\exsol{%
%$y = \frac{2(a-b)}{5} \, e^{-3x/2}+\frac{3 a+2 b}{5} \, e^x$
%}

\begin{exercise}%
    Find the solution to $y''-(\alpha+\beta) y' + \alpha \beta y=0$, $y(0) = a$, $y'(0)=b$, where $\alpha$, $\beta$, $a$, and $b$ are real numbers, and $\alpha \not= \beta$.
\end{exercise}
%\exsol{%
%$y =
%\frac{a \beta-b}{\beta-\alpha} e^{\alpha x} + 
%\frac{b-a \alpha}{\beta-\alpha} e^{\beta x}$
%}

\begin{exercise}%
    Write down an equation (guess) for which we have the solutions $e^x$ and $e^{2x}$.  Hint: Try an equation of the form $y''+Ay'+By = 0$ for constants $A$ and $B$, plug in both $e^x$ and $e^{2x}$ and solve for $A$ and $B$.
\end{exercise}
%\exsol{%
%$y''-3y'+2y = 0$
%}

\begin{exercise}%
    Construct an equation such that $y = C_1 e^{3x} + C_2 e^{-2x}$ is the general solution.
\end{exercise}
%\exsol{%
%$y'' -y'-6y=0$
%}

\begin{exercise}
    Give an example of a 2nd-order DE whose general solution is $y=c_1e^{-2t}+c_2e^{-4t}$.
\end{exercise}
%\comboSol{%
%}
%{%
%$y'' + 6y' + 8y = 0$
%}


Equations of the form $a x^2 y'' + b x y' + c y = 0$ are called \emph{Euler's equations\index{Euler's equation}} or \emph{Cauchy--Euler equations\index{Cauchy--Euler equation}}. They are solved by trying $y=x^r$ and solving for $r$ (assume that $x \geq 0$ for simplicity).

\begin{exercise} \label{sol:eulerex}
    Suppose that ${(b-a)}^2-4ac > 0$.
    \begin{tasks}
        \task Find a formula for the general solution of $a x^2 y'' + b x y' + c y = 0$.  Hint: Try $y=x^r$ and find a formula for $r$.
        \task What happens when ${(b-a)}^2-4ac = 0$ or ${(b-a)}^2-4ac < 0$?
    \end{tasks}
\end{exercise}
%\comboSol{%
%}
%{%
%a)~$r_1 = \frac{(a-b) + \sqrt{(b-a)^2 - 4ac}}{2a}$, $r_2 = \frac{(a-b) - \sqrt{(b-a)^2 - 4ac}}{2a}$, General Solution: $y = C_1x^{r_1} + C_2x^{r_2}$. \\
%b)~ Only one solution or complex solutions.
%}

We will revisit the case when ${(b-a)}^2-4ac < 0$ later.

\begin{exercise} \label{sol:eulerexln}
    Same equation as in \exerciseref{sol:eulerex}. Suppose ${(b-a)}^2-4ac = 0$.  Find a formula for the general solution of $a x^2 y'' + b x y' + c y = 0$.  Hint: Try $y=x^r \ln x$ for the second solution.
\end{exercise}
%\comboSol{%
%}
%{%
%$y = C_1x^{\frac{a-b}{2a}} + C_2x^{\frac{a-b}{2a}}ln(x)$
%}


%\setcounter{exercise}{100}



\end{document}