\documentclass{ximera}

\title{Practice for Autonomous Equations}

\input{../../preamble}

\begin{document}
\begin{abstract}
    Why?
\end{abstract}
\maketitle


\begin{exercise}
    Consider $x' = x^2$.
    \begin{tasks}
        \task Draw the phase diagram, find the critical points, and mark them asymptotically stable, semistable, or unstable.
        \task Sketch typical solutions of the equation.
        \task Find $\displaystyle \lim_{t\to \infty} x(t)$ for the solution with the initial condition
        $x(0) = -1$.
    \end{tasks}
\end{exercise}

%\comboSol{%
%}
%{%
%a)~ 0 is semistable \quad c)~ 0
%}


\begin{exercise}
    Consider $x' = \sin x$.
    \begin{tasks}
        \task Draw the phase diagram for $-4\pi \leq x \leq 4\pi$.  On this interval
        mark the critical points asymptotically stable, semistable, or unstable.
        \task Sketch typical solutions of the equation.
        \task Find $\displaystyle \lim_{t\to \infty} x(t)$ for the solution with the initial condition
        $x(0) = 1$.
    \end{tasks}
\end{exercise}

%%% TO-DO
%   Ximera has it's own way of dealing with solutions, so we should return to this later.
%%%
%\comboSol{%
%}
%{%
%a)~ $x = n\pi$ are critical points. $n$ odd is asymptotically stable, $n$ even is unstable. \quad c)~ $\pi$
%}

\begin{exercise}
    Let $x'=(x-1)(x-2)x^2$.
    \begin{tasks}
        \task Sketch the phase diagram and find critical
        points.
        \task Classify the critical points.
        \task If $x(0)=0.5$, then find $\displaystyle \lim_{t\to\infty} x(t)$.
    \end{tasks}
\end{exercise}


%%% TO-DO
%   Ximera has it's own way of dealing with solutions, so we should return to this later.
%%%
%\exsol{%
%a) 0, 1, 2 are critical points.
%\quad
%b) $x=0$ is unstable (semistable), $x=1$ is asymptotically stable, and $x=2$ is unstable.
%\quad
%c) 1
%}

\begin{exercise}
    Let $y' = (y-2)(y^2 + 1)(y+3)$. Sketch a phase diagram for this differential equation. Find and classify all critical points. If $y(0) = 0$, what will happen to the solution as $t \rightarrow \infty$?
\end{exercise}
%%% TO-DO
%   Ximera has it's own way of dealing with solutions, so we should return to this later.
%%%
%\comboSol{%
%}
%{%
%$y=2$ is unstable, $y=-3$ is asymptotically stable. Tends towards $-3$.
%}

\begin{exercise}
    Find and classify all equilibrium solutions for the differential equation $x' = (x-2)^2(x+1)(x+3)^3(x+2)$.
\end{exercise}

%%% TO-DO
%   Ximera has it's own way of dealing with solutions, so we should return to this later.
%%%
%\comboSol{%
%}
%{%
%$x=2$ is semistable, $x=-1$ is unstable, $x=-2$ is asymptotically stable, $x=-3$ is unstable.
%}

\begin{exercise}
    Let $y' = (y-3)(y+2)^2e^y$. Sketch a phase diagram for this differential equation. Find and classify all critical points. If $y(0) = 0$, what will happen to the solution as $t \rightarrow \infty$?
\end{exercise}
%%% TO-DO
%   Ximera has it's own way of dealing with solutions, so we should return to this later.
%%%
%\comboSol{%
%}
%{%
%$y=3$ is unstable, $y=-2$ is semistable. Tends towards $-2$.
%}

\begin{exercise}
    Consider the DE $\dfrac{dy}{dt}=y^5-3y^4+3y^3-y^2$. %1.7
    Find and classify all equilibrium solutions of this DE. Then sketch a representative selection of solution curves.
\end{exercise}
%%% TO-DO
%   Ximera has it's own way of dealing with solutions, so we should return to this later.
%%%
%\comboSol{%
%}
%{%
%$y=1$ is unstable, $y=0$ is semistable.
%}

\begin{exercise}
    Let $x'=e^{-x}$.
    \begin{tasks}(2)
        \task Find and classify all critical points.
        \task Find $\displaystyle \lim_{t\to\infty} x(t)$ given any initial condition.
    \end{tasks}
\end{exercise}

%%% TO-DO
%   Ximera has it's own way of dealing with solutions, so we should return to this later.
%%%
%\exsol{%
%a) There are no critical points.
%\quad
%b) $\infty$
%}


\begin{exercise}
    Suppose $f(x)$ is positive for $0 < x < 1$, it is zero when $x=0$ and $x=1$, and it is negative for all other $x$.
    \begin{tasks}
        \task Draw the phase diagram for $x' = f(x)$, find the critical points, and mark them asymptotically stable, semistable, or unstable.
        \task Sketch typical solutions of the equation.
        \task Find $\displaystyle \lim_{t\to \infty} x(t)$ for the solution with the initial condition $x(0) = 0.5$.
    \end{tasks}
\end{exercise}
%%% TO-DO
%   Ximera has it's own way of dealing with solutions, so we should return to this later.
%%%
%\comboSol{%
%}
%{%
%a)~ $x=1$ is asymptotically stable, $x=0$ is unstable. \quad c)~ Tends to 1.
%}

\begin{exercise}
    Suppose $\frac{dx}{dt} = (x-\alpha)(x-\beta)$ for two numbers $\alpha < \beta$.
    \begin{tasks}
        \task Find the critical points, and classify them.
    \end{tasks}
    For b), c), d), find $\displaystyle \lim_{t\to\infty} x(t)$ based on the phase diagram.
    \begin{tasks}[resume](3)
        \task $x(0) < \alpha$,
        \task $\alpha < x(0) < \beta$,
        \task $\beta < x(0)$.
    \end{tasks}
\end{exercise}

%%% TO-DO
%   Ximera has it's own way of dealing with solutions, so we should return to this later.
%%%
%\exsol{%
%a) $\alpha$ is a stable critical point, $\beta$ is an unstable one.
%\quad
%b) $\alpha$, \quad c) $\alpha$, \quad d) $\infty$ or DNE\@.
%}

\begin{exercise}
    A disease is spreading through the country.  Let $x$ be the number of people infected.  Let the constant $S$ be the number of people susceptible to infection.  The infection rate $\frac{dx}{dt}$ is proportional to the product of already infected people, $x$, and the number of susceptible but uninfected people, $S-x$.
    \begin{tasks}
        \task Write down the differential equation.
        \task Supposing $x(0) > 0$, that is, some people are infected at time $t=0$, what is $\displaystyle \lim_{t\to\infty} x(t)$.
        \task Does the solution to part b) agree with your intuition?  Why or why not?
    \end{tasks}
\end{exercise}
%%% TO-DO
%   Ximera has it's own way of dealing with solutions, so we should return to this later.
%%%
%\comboSol{%
%}
%{%
%a)~ $\frac{dx}{dt} = k(x)(S - x)$ \quad b)~ S \qquad c)~ Yes, everyone gets it.
%}


%\setcounter{exercise}{100}






\end{document}