\documentclass{ximera}

\title{Practice for Modeling Parameters}

\input{../../preamble}

\begin{document}
\begin{abstract}
    Why?
\end{abstract}
\maketitle


\begin{exercise}
    Bob is getting coffee from a restaurant and knows that the temperature of the coffee will follow Newton's Law of Cooling, which says that
    \begin{equation*}
        \frac{dT}{dt} = k(T_0 - T)
    \end{equation*}
    where $T_0$ is the ambient temperature and $k$ is a constant depending on the object and geometry. His car is held at a constant 20$^\circ$ C, and when he receives the coffee, he measures the temperature to be 90$^\circ$ C. Two minutes later, the temperature is 81$^\circ$C.
    
    Use these two points of data to determine the value of $k$ for this coffee. [Enter your answer to 5 decimal digits] $\answer{0.06681}$
    
    \begin{problem}
        Bob only wants to drink his coffee once it reaches 65$^\circ$ C. How long does he have to wait for this to happen? [Enter your answer to 2 decimal digits] $\answer{6.42}$ minutes
        \begin{problem}
            If the coffee is too cold for Bob's taste once it reaches 35$^\circ$ C, how long is the acceptable window for Bob to drink his coffee? [Enter as a whole number] $\answer{16}$ minute(s)
        \end{problem}
    \end{problem} 
\end{exercise}
%\comboSol{%
%}
%{%
%a)~ $\approx$ 0.06681 \quad b)~$\approx$ 6.42 min \quad c)~ $\approx$ 16 min.
%}

\begin{exercise}
    Assume that a falling object has a velocity (in meters per second) that obeys the differential equation
    \begin{equation*}
        \frac{dv}{dt} = 9.8 - \alpha v
    \end{equation*}
    where $\alpha$ represents the drag coefficient of the object.
    
    Solve this differential equation with initial condition $v(0) = 0$ to get a solution that depends on $\alpha$. 
    \[
        v(t) = \answer{\frac{9.8}{\alpha}(1 - e^{-\alpha t})}
    \]
    \begin{problem}
        Assume that you drop an object from a height of 10 meters and it hits the ground after $3$ seconds. What is the value of $\alpha$ here? (Note: You solved for $v(t)$ in the previous part, and now you need to get to position. What does that require?) [Enter answer to two decimal digits] $\answer{2.94}$
        \begin{problem}
            Assume that a second object hits the ground in 6 seconds. How does this change the value of $\alpha$? [Enter to two decimal digits] $\answer{5.88}$
        \end{problem}
    \end{problem}
\end{exercise}
%\comboSol{%
%}
%{%
%a)~ $v(t) = \frac{9.8}{\alpha}(1 - e^{-\alpha t})$ \quad b)~2.94 \quad c)~5.88
%}

\begin{exercise}
    A restaurant is trying to analyze the to-go coffee cups that it uses in order to best serve their customers. They assume that the coffee follows Newton's Law of Cooling and place it in a room with ambient temperature 22$^\circ$ C. They record the following data for the temperature of the coffee as a function of time.
    \begin{table}[h!!]
        \centering
        \begin{tabular}{|c|c|}\hline
             \textbf{t} (min)& \textbf{T} ($^\circ$ C)  \\ \hline
            0 & 80 \\
             0.5 & 74 \\
             1.1 & 67 \\
             1.7 & 61 \\
             2.3 & 56 \\ \hline
        \end{tabular}
    \end{table}
    Use code to determine the best-fit value of $k$ for this data. [Enter to 4 decimal digits] $\answer{0.2317}$
    \begin{problem}
        The restaurant determines that to avoid any potential legal issues, the coffee can be no warmer than 60 $^\circ$C when it is served. If the coffee comes out of the machine at 90 $^\circ C$, how long do they have to wait before they can serve the coffee? [Enter to 2 decimal digits] $\answer{2.51}$ minute(s).
    \end{problem}
\end{exercise}
%\comboSol{%
%}
%{%
%a)~ $k \approx 0.2317$ \quad b)~$\approx 2.51$ min
%}

\newpage

\begin{exercise}\label{ex:ModelingParamVel}%
    Assume that an object falling has a velocity that follows the differential equation 
    \begin{equation*}
        \frac{dv}{dt} = 9.8 - \alpha v^2
    \end{equation*}
    where the velocity is in \unitfrac{m}{s} and $\alpha$ represents the drag coefficient. During a physics experiment, a student measures data for the velocity of a falling object over time given in the table below.
    
    Use this data (and code) to estimate the value of $\alpha$ for this object.  [Enter your answer to 4 decimal digits] $\answer{0.1244}$.
\end{exercise} 
%\exsol{%
%$\alpha = .1244$. The alpha value used before noise was added to the data is $0.124$, so very close, but not identically the same.%
%}



\begin{exercise}\label{ex:ModelingParamPop}
    Assume that a species of fish in a lake has a population that is modeled by the differential equation
    \begin{equation*}
        \frac{dP}{dt} = \frac{1}{100}rP(K - P) - \alpha 
    \end{equation*}
    where $r$, $K$, and $\alpha$ are parameters, $r$ representing the growth rate, $K$ the carrying capacity, and $\alpha$ the harvesting rate, and the population $P$ is in thousands., with $t$ given in years. From previous studies, you know that the best value of $r$ is $3.12$. After studying the population over a period of time, you get the following data:
    
    %\begin{minipage}{0.49\textwidth}
    %\centering
        \begin{tabular}{|c|c|}\hline
             \textbf{t} (s)& \textbf{v} (m/s)  \\ \hline
            0 & 0 \\
        	0.1 &1.0 \\
        	0.2 &  1.9\\
        	0.4 & 3.6 \\
        	0.6 & 5.2 \\
        	0.9 & 6.8 \\
        	1.1 & 7.4 \\
        	1.3 & 7.9 \\
        	1.5 & 8.2 \\
        	1.8 & 8.5\\
        	2.1 & 8.8  \\ \hline
        \end{tabular}
    %    \captionof{table}{Data for \exerciseref{ex:ModelingParamVel}.}
    %\end{minipage}%
    
    %\begin{minipage}{0.49\textwidth}
    %    \centering
        \begin{tabular}{|c|c|}\hline
             \textbf{t} (d)& \textbf{P} (thousands)  \\ \hline
            0 & 50 \\
           7  & 60 \\
            14 & 70 \\
            28 & 97 \\
            37 & 117 \\
             50 & 148 \\
            78 & 220 \\
            100 & 268
             \\ \hline
        \end{tabular}
    %    \captionof{table}{Data for \exerciseref{ex:ModelingParamPop}.}
    %\end{minipage}    
    
    Your friend tells you that in a previous study, he found that the value of $K$ for this particular lake is $450$. Use code to determine the best value of $\alpha$ for this situation. Note that the equation expects $t$ in years, but the data is given in days. Search for $\alpha$ in the range $(0, 400)$. [Enter your answer to 2 decimal places] $\alpha = \answer{337.55}$.
    \begin{feedback}[correct]
        That answer doesn't look great. Plot the solution with these parameters along with the data and compare them. 
    \end{feedback}
    \begin{problem}
        The fit does not look great, so maybe your friend's value was not quite right. Run code to find best values for $K$ and $\alpha$ simultaneously. Use the range $(0, 400)$ for both $\alpha$ and $K$. [Enter your answer to two decimal places]\\
        $K = \answer{350.45}$\\
        $\alpha = \answer{3.79}$
    \end{problem}

\end{exercise}
%\comboSol{%
%}
%{%
%a)~ $\alpha = 337.55$ \quad c)~ $\alpha = 6.53$, $K = 350.45$. (The base data was $K = 350$, $\alpha = 3.79$.)
%}



\end{document}