\documentclass{ximera}

\title{Practice for Euler Method}

\input{../../preamble}

\begin{document}
\begin{abstract}
    Why?
\end{abstract}
\maketitle


\begin{exercise}
    Consider $\dfrac{dx}{dt} = {(2t-x)}^2$, $x(0)=2$.  Use Euler's method with step size $h=0.5$ to approximate $x(1)$.
\end{exercise}
%\comboSol{%
%}
%{%
%17/2
%}

\begin{exercise}
    Consider the differential equation $\frac{dy}{dt} = t^2 - 3y + 1$ with $y(1) = 4$. Approximate the solution at $t=3$ using Euler's method with a step size of $h=1$ and $h=0.5$. Compare these values with the actual solution at $t=3$. 
\end{exercise}
%\comboSol{%
%}
%{%
%$h=1$: 17, $h=0.5$: 85/32, Actual: $\frac{52}{27} + \frac{94}{27}e^{-6} \approx 1.9346$
%}

\begin{exercise}
    Consider the differential equation $\frac{dy}{dt} = 2ty + y^2$ with $y(0) = 1$. Approximate the solution at $t=2$ using Euler's method with a step size of $h=1$ and $h=0.5$.
\end{exercise}
%\comboSol{%
%}
%{%
%$h=1$: 10, $h=0.5$: $y(2) \approx 108.56$
%}

\begin{samepage}
    \begin{exercise}
        Consider $\dfrac{dx}{dt} = t-x$, $x(0)=1$.
        \begin{tasks}
            \task Use Euler's method with step sizes $h = 1, \nicefrac{1}{2}, \nicefrac{1}{4}, \nicefrac{1}{8}$ to approximate $x(1)$. 
            \task Solve the equation exactly.
            \task Describe what happens to the errors for each $h$ you used.  That is, find the factor by which the error changed each time you halved the interval.
        \end{tasks}
    \end{exercise}
\end{samepage}
%\comboSol{%
%}
%{%
%a)~ $h=1$, 0. $h=\nicefrac{1}{2}$, 0.5. $h = \nicefrac{1}{4}$,$ \frac{81}{128}$. $h = \nicefrac{1}{8}$ $x(1) \approx .6872$. \\
%b)~$x(t) = t-1+2e^{-t}$\\
%c)~Error decreases by factors of 0.32, 0.4365, 0.4723.
%}

\begin{exercise}%
    Let $x' = \sin(xt)$, and $x(0)=1$. Approximate $x(1)$ using Euler's method with step sizes 1, 0.5, 0.25. Use a calculator and compute up to 4 decimal digits.
\end{exercise}
%\exsol{%
%Approximately: 1.0000, 1.2397, 1.3829
%}


\begin{exercise}
    Approximate the value of $e$ by looking at the initial value problem $y'=y$ with $y(0)=1$ and approximating $y(1)$ using Euler's method with a step size of $0.2$.
\end{exercise}
%\comboSol{%
%}
%{%
%2.4883
%}

\begin{exercise}%
    Let $x' = 2t$, and $x(0)=0$.
    \begin{tasks}
        \task Approximate $x(4)$ using Euler's method with step sizes 4, 2, and 1.
        \task Solve exactly, and compute the errors.
        \task Compute the factor by which the errors changed.
    \end{tasks}
\end{exercise}
%\exsol{%
%a) 0, 8, 12
%\quad
%b) $x(4) = 16$, so errors are: 16, 8, 4.
%\quad
%c) Factors are 0.5, 0.5, 0.5.
%}

\begin{samepage}
    \begin{exercise}%
        Let $x' = x e^{xt+1}$, and $x(0)=0$.
        \begin{tasks}
            \task Approximate $x(4)$ using Euler's method with step sizes 4, 2, and 1.
            \task Guess an exact solution based on part a) and compute the errors.
        \end{tasks}
    \end{exercise}
\end{samepage}
%\exsol{%
%a) 0, 0, 0
%\quad
%b) $x=0$ is a solution so errors are: 0, 0, 0.
%}


\begin{exercise}
    Example of numerical instability: Take $y' = -5y$, $y(0) = 1$.  We know that the solution should decay to zero as $x$ grows. Using Euler's method, start with $h=1$ and compute $y_1, y_2, y_3, y_4$ to try to approximate $y(4)$.  What happened? Now halve the interval.  Keep halving the interval and approximating $y(4)$ until the numbers you are getting start to stabilize (that is, until they start going towards zero). Note: You might want to use a calculator.
\end{exercise}
%\comboSol{%
%}
%{%
%For $h=1$, $y_4 = 256$. $h=0.5$, $y_4 = 25.6289$, $h=0.25$ goes to zero, but oscillates. $h=0.125$ just goes to zero.
%}

There is a simple way to improve Euler's method to make it a second order method by doing just one extra step. Consider $\frac{dy}{dx}=f(x,y)$, $y(x_0) = y_0$, and a step size $h$. What we do is to pretend we compute the next step as in Euler, that is, we start with $(x_i,y_i)$, we compute a slope $k_1 = f(x_i,y_i)$, and then look at the point $(x_i+h,y_i + k_1h)$. Instead of letting our new point be $(x_i+h,y_i + k_1h)$, we compute the slope at that point, call it $k_2$, and then take the average of $k_1$ and $k_2$, hoping that the average is going to be closer to the actual slope on the interval from $x_i$ to $x_i+h$.  And we are correct, if we halve the step, the error should go down by a factor of $2^2 = 4$. To summarize, the setup is the same as for regular Euler, except the computation of $y_{i+1}$ and $x_{i+1}$.
\begin{align*}
    & k_1 = f(x_i,y_i) , & & 
    x_{i+1} = x_i + h , \\
    & k_2 = f(x_i + h,y_i + k_1h) ,
    & & y_{i+1} = y_i + \frac{k_1+k_2}{2}\,h .
\end{align*}


\begin{exercise}%
    Consider $\dfrac{dy}{dx} = x+y$, $y(0)=1$.
    \begin{tasks}
        \task Use the improved Euler's method (see above) with step sizes $h=\nicefrac{1}{4}$ and $h=\nicefrac{1}{8}$ to approximate $y(1)$.
        \task Use Euler's method with $h=\nicefrac{1}{4}$ and $h=\nicefrac{1}{8}$.
        \task Solve exactly, find the exact value of $y(1)$.
        \task Compute the errors, and the factors by which the errors changed.
    \end{tasks}
\end{exercise}
%\exsol{%
%a) Improved Euler: $y(1) \approx 3.3897$ for $h=\nicefrac{1}{4}$, 
%$y(1) \approx 3.4237$ for $h=\nicefrac{1}{8}$, 
%\quad
%b) Standard Euler: $y(1) \approx 2.8828$ for $h=\nicefrac{1}{4}$, 
%$y(1) \approx 3.1316$ for $h=\nicefrac{1}{8}$, 
%\quad
%c) $y = 2e^x-x-1$, so $y(2)$ is approximately $3.4366$.
%\quad
%d) Approximate errors for improved Euler:
%$0.046852$ for $h=\nicefrac{1}{4}$, 
%and
%$0.012881$ for $h=\nicefrac{1}{8}$. 
%\quad
%For standard Euler:
%$0.55375$ for $h=\nicefrac{1}{4}$, 
%and
%$0.30499$ for $h=\nicefrac{1}{8}$. 
%\quad
%Factor is approximately $0.27$ for improved Euler, and
%$0.55$ for standard Euler.
%}


The simplest method used in practice is the \emph{Runge--Kutta method}. Consider $\frac{dy}{dx}=f(x,y)$, $y(x_0) = y_0$, and a step size $h$.  Everything is the same as in Euler's method, except the computation of $y_{i+1}$ and $x_{i+1}$.
\begin{align*}
    & k_1 = f(x_i,y_i) , & & \\
    & k_2 = f\bigl(x_i + \nicefrac{h}{2},y_i + k_1 (\nicefrac{h}{2})\bigr) ,
    & & 
    x_{i+1} = x_i + h , \\
    & k_3 = f\bigl(x_i + \nicefrac{h}{2},y_i + k_2 (\nicefrac{h}{2})\bigr) ,
    & &
    y_{i+1} = y_i + \frac{k_1 + 2k_2 + 2k_3 + k_4}{6}\,h ,  \\
    & k_4 = f(x_i + h,y_i + k_3 h) .
\end{align*}


\begin{exercise}
    Consider $\dfrac{dy}{dx} = yx^2$, $y(0)=1$.
    \begin{tasks}
        \task Use Runge--Kutta (see above) with step sizes $h=1$ and $h=\nicefrac{1}{2}$ to approximate $y(1)$.
        \task Use Euler's method with $h=1$ and $h=\nicefrac{1}{2}$.
        \task Solve exactly, find the exact value of $y(1)$, and compare.
    \end{tasks}
\end{exercise}
%\comboSol{%
%}
%{%
%a)~ $h=1$ gives 1.3906. $h=0.5$ gives 1.3953.
%b)~ $h=1$ gives 1. $h=0.5$ gives 1.125.
%c)~ Exact: 1.3956. Runge--Kutta matches three decimal places at $h=0.5$. Euler needs to go to $h=0.005$ to get the first two decimals right.
%}
%
%\setcounter{exercise}{100}


\end{document}