\documentclass{ximera}

\title{Practice for Higher Eval Problems}

\input{../../preamble}

\begin{document}
\begin{abstract}
Why?
\end{abstract}
\maketitle


\begin{exercise}
    Suppose you have a beam of length 5 with free ends.  Let $y$ be the transverse deviation of the beam at position $x$ on the beam ($0 < x < 5$). You know that the constants are such that this satisfies the equation $y_{tt} + 4 y_{xxxx} = 0$. Suppose you know that the initial shape of the beam is the graph of $x(5-x)$, and the initial velocity is uniformly equal to 2 (same for each $x$) in the positive $y$ direction.  Set up the equation together with the boundary and initial conditions.  Just set up, do not solve.
\end{exercise}

\begin{exercise}
    Suppose you have a beam of length 5 with one end free and one end fixed (the fixed end is at $x=5$). Let $u$ be the longitudinal deviation of the beam at position $x$ on the beam ($0 < x < 5$). You know that the constants are such that this satisfies the equation $u_{tt} = 4 u_{xx}$. Suppose you know that the initial displacement of the beam is $\frac{x-5}{50}$, and the initial velocity is $\frac{-(x-5)}{100}$ in the positive $u$ direction.  Set up the equation together with the boundary and initial conditions.  Just set up, do not solve.
\end{exercise}

\begin{exercise}
    Suppose the beam is $L$ units long, everything else kept the same as in \eqref{appeig:beameq}.  What is the equation and the series solution?
\end{exercise}

\begin{exercise}
    Suppose you have 
    \begin{equation*}
        \begin{aligned}
            & a^4 y_{xxxx} + y_{tt} = 0 \quad (0 < x < 1, t > 0) , \\
            & y(0,t) = y_{xx}(0,t) = 0,\\
            & y(1,t) = y_{xx}(1,t) = 0 ,\\
            & y(x,0) = f(x), \quad y_{t}(x,0) = g(x) .
        \end{aligned}
    \end{equation*}
    That is, you have also an initial velocity.  Find a series solution.  Hint: Use the same idea as we did for the wave equation.
\end{exercise}

%\setcounter{exercise}{100}

\begin{exercise}
    Suppose you have a beam of length 1 with hinged ends.  Let $y$ be the transverse deviation of the beam at position $x$ on the beam ($0 < x < 1$). You know that the constants are such that this satisfies the equation $y_{tt} + 4 y_{xxxx} = 0$.   Suppose you know that the initial shape of the beam is the graph of $\sin (\pi x)$, and the initial velocity is 0.  Solve for $y$.
\end{exercise}
%\exsol{%
%$y(x,t) = \sin(\pi x) \cos (2 \pi^2 t)$
%}

\begin{exercise}
    Suppose you have a beam of length 10 with two fixed ends.  Let $y$ be the transverse deviation of the beam at position $x$ on the beam ($0 < x < 10$). You know that the constants are such that this satisfies the equation $y_{tt} + 9 y_{xxxx} = 0$.   Suppose you know that the initial shape of the beam is the graph of $\sin(\pi x)$, and the initial velocity is uniformly equal to $x(10-x)$. Set up the equation together with the boundary and initial conditions.  Just set up, do not solve.
\end{exercise}
%\exsol{%
%$9 y_{xxxx} + y_{tt} = 0 \quad (0 < x < 10, t > 0)$, \quad
%$y(0,t) = y_{x}(0,t) = 0$, \quad
%$y(10,t) = y_{x}(10,t) = 0$, \quad
%$y(x,0) = \sin(\pi x), \quad y_{t}(x,0) = x(10-x)$.
%}


\end{document}