\documentclass{ximera}

\title{Practice for Steady Per Solutions}

\input{../../preamble}

\begin{document}
\begin{abstract}
Why?
\end{abstract}
\maketitle


\begin{exercise} \label{sps:sinforceexr}
    Suppose that the forcing function for the vibrating string is $F_0 \sin (\omega t)$.  Derive the particular solution $y_p$.
\end{exercise}

\begin{exercise}
    Take the forced vibrating string. Suppose that $L=1$, $a=1$.  Suppose that the forcing function is the square wave that is 1 on the interval $0 < x < 1$ and $-1$ on the interval $-1 < x< 0$. Find the particular solution.  Hint: You may want to use result of exercise \ref{sps:sinforceexr}.
\end{exercise}

\begin{exercise}
    The units are cgs (centimeters-grams-seconds). For $k=0.005$, $\omega = 1.991 \times {10}^{-7}$, $A_0 = 20$. Find the depth at which the temperature variation is half ($\pm 10$ degrees) of what it is on the surface.
\end{exercise}

\begin{exercise}
    Derive the solution for underground temperature oscillation without assuming that $T_0 = 0$.
\end{exercise}

%\setcounter{exercise}{100}

\begin{exercise}
    Take the forced vibrating string. Suppose that $L=1$, $a=1$.  Suppose that the forcing function is a sawtooth, that is $\lvert x \rvert -\frac{1}{2}$ on $-1 < x < 1$ extended periodically. Find the particular solution.
\end{exercise}
%\exsol{%
%$y_p(x,t) =
%\sum\limits_{\substack{n=1 \\ n \text{ odd}}}^\infty
%\frac{-4}{n^4 \pi^4}
%\left(
%\cos(n \pi x ) -
%\frac{\cos ( n \pi ) - 1}%
%{\sin( n \pi)}
%\sin( n \pi x)
%-1
%\right)
%\cos (n \pi t) .
%$
%}

\begin{exercise}
    The units are cgs (centimeters-grams-seconds). For $k=0.01$, $\omega = 1.991 \times {10}^{-7}$, $A_0 = 25$. Find the depth at which the summer is again the hottest point.
\end{exercise}
%\exsol{%
%%$x \sqrt{\frac{\omega}{2k}} = 2\pi$
%%$x = 2\pi\sqrt{\frac{2k}{\omega}}$
%%$x = 2\pi\sqrt{\frac{0.02}{1.991 x 10^-7}}$
%Approximately 1991 centimeters
%}


\end{document}