\documentclass{ximera}

\title{Practice for Higher Order ODEs}

\input{../../preamble}

\begin{document}
\begin{abstract}
    Why?
\end{abstract}
\maketitle

 

\begin{exercise}
    Find the general solution for $y''' - y'' + y' - y = 0$.
\end{exercise}
%\comboSol{%
%}
%{%
%$y = C_1e^t + C_2\cos(t) + C_3\sin(t)$
%}

\begin{exercise}%
    Find the general solution of $y^{(5)}-y^{(4)}=0$.
\end{exercise}
%\exsol{%
%$y=C_1 e^x +C_2 x^3 + C_3 x^2 +C_4 x + C_5$
%}

\begin{exercise}
    Find the general solution for $y^{(4)} - 5 y''' + 6 y'' = 0$.
\end{exercise}
%\comboSol{%
%}
%{%
%$y = C_1e^{3t} + C_2e^{2t} + C_3 + C_4t$
%}

\begin{exercise}
    Find the general solution for $y''' + 2 y'' + 2 y' = 0$.
\end{exercise}
%\comboSol{%
%}
%{%
%$y = C_1 + C_2e^{-t}\cos(t) + C_3e^{-t}\sin(t)$
%}

\begin{exercise}
    Suppose the characteristic equation for an ODE is ${(r-1)}^2{(r-2)}^2 = 0$.
    \begin{tasks}
        \task Find such a differential equation.
        \task Find its general solution.
    \end{tasks}
\end{exercise}
%\comboSol{%
%}
%{%
%a)~ $y^{(4)} - 6y''' + 13y'' - 12y' + 4y = 0$ \quad b)~$y = C_1e^t + C_2te^t + C_3e^{2t} + C_4te^{2t}$
%}

\begin{exercise} \label{hol:eqfromsolex}
    Suppose that a fourth order equation has a solution $y = 2 e^{4x} x \cos x$.  
    \begin{tasks}
        \task Find such an equation.
        \task Find the initial conditions that the given solution satisfies.
    \end{tasks}
\end{exercise}
%\comboSol{%
%}
%{%
%a)~$y^{(4)} + 16y''' + 20y'' + 272y' + 289y = 0$ \quad b)~$y(0) = 0$, $y'(0) = 2$, $y''(0) = 16$, $y'''(0) = 90$
%}

\begin{exercise}
    Suppose that the characteristic equation of a third order differential equation has roots $\pm 2i$ and 3.
    \begin{tasks}
        \task What is the characteristic equation?
        \task Find the corresponding differential equation.
        \task Find the general solution.
    \end{tasks}
\end{exercise}
%\exsol{%
%a) $r^3-3r^2+4r-12 = 0$
%\quad
%b) $y'''-3y''+4y'-12y = 0$
%\quad
%c) $y = C_1 e^{3x} + C_2 \sin(2x) + C_3 \cos(2x)$
%}

\begin{exercise}
    Find the general solution for the equation of \exerciseref{hol:eqfromsolex}.
\end{exercise}
%\comboSol{%
%}
%{%
%$y(x) = C_1e^{4x}\cos(x) + C_2xe^{4x}\cos(x) + C_3e^{4x}\sin(x) + C_4xe^{4x}\sin(x)$
%}

\begin{exercise}%
    Find the general solution of
    \begin{equation*}
        y^{(4)} - y''' - 5y'' - 23y' - 20y = 0.
    \end{equation*}
\end{exercise}
%\exsol{%
%$y(x) = C_1e^{4x} + C_2e^{-x} + C_3e^{-x}\cos(2x) + C_4e^{-x}\sin(2x)$
%}

\begin{exercise}
    Find the general solution of
    \begin{equation*}
        y''' - 6y'' + 13y' - 10y = 4e^x + 5e^{3x} - 20.
    \end{equation*}
\end{exercise}
%\comboSol{%
%}
%{%
%$y(x) = C_1 e^{2x} + C_2e^{2x}\cos(x) + C_3e^{2x}\sin(x) - 2e^x + \frac{1}{16}e^{3x} - 2$
%}


\begin{exercise}
    Find the general solution of
    \begin{equation*}
        y''' - 3y' + 2y = 2e^{x} - e^{3x}.
    \end{equation*}
\end{exercise}
%\comboSol{%
%}
%{%
%$y(x) = (C_1 + C_2x)e^{x} + C_3e^{-2x} + \frac{1}{3}x^2e^x - \frac{1}{20}e^{3x}$
%}


\begin{exercise}
    Find the general solution of
    \begin{equation*}
        y''' + 2y'' + y' + 2y = 3\cos(x) + x.
    \end{equation*}
\end{exercise}
%\comboSol{%
%}
%{%
%$y(x) =  C_1e^{-2x} + C_2cos(x) + C_3sin(x) - \frac{3}{10}x\cos(x) + \frac{3}{5}x\sin(x) + \frac{1}{2}x - \frac{1}{4}$
%}

\begin{exercise}
    Find the general solution of 
    \begin{equation*}
        y^{(4)}  + 2y'' + y = 4x\cos(x) - e^{3x} + 1
    \end{equation*}
    Hint: Remember, the guess needs to make sure that no terms in it solve the homogeneous equation.
\end{exercise}
%\comboSol{%
%}
%{%
%$y(x) = (C_1 + C_2x)cos(x) + (C_3 + C_4x)sin(x) - \frac{1}{6}x^3\cos(x) + \frac{1}{2}x^2\sin(x) - \frac{1}{100}e^{3x} + 1$
%}


\begin{exercise}
    Show that $y=\cos(2t)$ is a solution to $y^{(4)}+2y'''+9y''+8y'+20y=0$. %2.7
    This tells us something about the factorization of the characteristic polynomial of this DE. Factor the characteristic polynomial completely, and solve the DE.
\end{exercise}
%\comboSol{%
%}
%{%
%$y(t) = C_1\cos(2t) + C_2\sin(2t) + C_3e^{-t}\cos(2t) + C_3e^{-t}\sin(2t)$
%}

\begin{exercise}
    Consider 
    \begin{equation}
        y'''-y''-8y'+12y=0. \label{eq:2_7_Exercise1}
    \end{equation}
    \begin{tasks}
        \task Show that $y=e^{2t}$ is a solution of \eqref{eq:2_7_Exercise1}.
        \task Find the general solution to \eqref{eq:2_7_Exercise1}.
        \task Solve $y'''-y''-8y'+12y=e^{2t}$.
    \end{tasks}
\end{exercise}
%\comboSol{%
%}
%{%
%b)~ $y(t) = C_1e^{2t} + C_2te^{2t} + C_3e^{-3t}$ \\
%c)~$y(t) = C_1e^{2t} + C_2te^{2t} + C_3e^{-3t} + \frac{1}{10}t^2e^{2t}$
%}

\begin{exercise}
    Let $f(x) = e^x - \cos x$, $g(x) = e^x + \cos x$, and $h(x) = \cos x$. Are $f(x)$, $g(x)$, and $h(x)$ linearly independent?  If so, show it, if not, find a linear combination that works.
\end{exercise}
%\comboSol{%
%}
%{%
%No
%}

\begin{exercise}
    Let $f(x) = 0$, $g(x) = \cos x$, and $h(x) = \sin x$. Are $f(x)$, $g(x)$, and $h(x)$ linearly independent?  If so, show it, if not, find a linear combination that works. 
\end{exercise}
%\comboSol{%
%}
%{%
%No
%}

\begin{exercise}%
    Are $e^{x}$, $e^{x+1}$, $e^{2x}$, $\sin(x)$ linearly independent? If so, show it, if not find a linear combination that works.
\end{exercise}
%\exsol{%
%No.  $e^1 e^x -  e^{x+1} = 0$.
%}

\begin{exercise}
    Are $x$, $x^2$, and $x^4$ linearly independent?  If so, show it, if not, find a linear combination that works.
\end{exercise}
%\comboSol{%
%}
%{%
%Yes. (Hint: How many roots can a polynomial have?)
%}

\begin{exercise}
    Are $e^x$, $xe^x$, and $x^2e^x$ linearly independent?  If so, show it, if not, find a linear combination that works.
\end{exercise}
%\comboSol{%
%}
%{%
%Yes. Factor out an $e^x$ first.
%}

\begin{exercise}%
    Are $\sin(x)$, $x$, $x\sin(x)$ linearly independent? If so, show it, if not find a linear combination that works.
\end{exercise}
%\exsol{%
%Yes.  (Hint: First note that $\sin(x)$ is bounded.  Then note that
%$x$ and $x\sin(x)$ cannot be multiples of each other.)
%}

\begin{exercise}
    Show that $\{e^t, te^t, e^{-t}, te^{-t}\}$ is a linearly independent set.
\end{exercise}
%\comboSol{%
%}
%{%
%Hint: Try to find the $t$ value when they match, and determine that it can't be all $t$. 
%}

\begin{exercise}%
    Solve $1001y'''+3.2y''+\pi y'-\sqrt{4} y = 0$, $y(0)=0$, $y'(0) = 0$, $y''(0) = 0$.
\end{exercise}
%\exsol{%
%$y=0$
%}

\begin{exercise}
    Could $y=t^2\cos t$ be a solution of a homogeneous DE with constant real coefficients? %2.7
    If so, give the minimum possible order of such a DE, and state which functions must also be solutions. If not, explain why this is impossible.
\end{exercise}
%\comboSol{%
%}
%{%
%Yes. Order 6. $\cos(t)$, $t\cos(t)$, $\sin(t)$, $t\sin(t)$, $t^2\sin(t)$
%}

\begin{exercise}
    Find a linear DE with constant real coefficients whose general solution is $$y=c_1e^{2t}+ c_2e^{-t}\cos(4t) + c_3e^{-t}\sin(2t),$$ %2.7
    or explain why there is no such thing.
\end{exercise}
%\comboSol{%
%}
%{%
%Not possible.
%}

\begin{exercise}
    Find an equation such that $y=xe^{-2x}\sin(3x)$ is a solution.
\end{exercise}
%\comboSol{%
%}
%{%
%$y^{(4)} + 8y''' + 42y'' + 104y' + 169y = 0$
%}

\begin{exercise}%
    Find an equation of minimal order such that $y=\cos(x)$, $y=\sin(x)$, $y=e^x$ are solutions.
\end{exercise}
%\exsol{%
%$y'''-y''+y'-y=0$
%}

\begin{exercise}
    Find an equation of minimal order such that $y=\cos(x)$, $y=\sin(2x)$, $y=e^{3x}$ are solutions.
\end{exercise}
%\comboSol{%
%}
%{%
%$y^{(5)} - 3y^{(4)} + 5y^{(3)} - 15y'' + 4y' - 12y = 0$
%}

\begin{exercise}
    Find a homogeneous DE with general solution %2.7
    $$y=c_1e^t+ c_2e^{-t}+ c_3\cos t+ c_4\sin t + c_5te^t+ c_6te^{-t}+ c_7t\cos t+ c_8t\sin t.$$
\end{exercise}
%\comboSol{%
%}
%{%
%$y^{(8)} - 2y^{(4)} + y = 0$
%}



%\setcounter{exercise}{100}


\end{document}