\documentclass{ximera}

\title{Practice for Convolution}

\input{../../preamble}

\begin{document}
\begin{abstract}
Why?
\end{abstract}
\maketitle


\begin{exercise}
    Using the Heaviside function write down the piecewise function that is 0 for $t < 0$, $t^2$ for $t$ in $[0,1]$ and $t$ for $t > 1$.
\end{exercise}

\begin{exercise}%
    Using the Heaviside function $u(t)$, write down the function
    \begin{equation*}
        f(t) =
        \begin{cases}
            0 & \text{if } \; \phantom{1 \leq {}} t < 1  , \\
            t-1 & \text{if } \; 1 \leq t < 2 , \\
            1 & \text{if } \; 2 \leq t .
        \end{cases}
    \end{equation*}
\end{exercise}
%\exsol{%
%$f(t) =
%(t-1)\bigl(u(t-1) - u(t-2)\bigr) + 
%u(t-2)$
%}

\begin{exercise}
    Using the Laplace transform solve
    \begin{equation*}
        m x'' + c x' + k x = 0 , \quad x(0) = a, \quad x'(0) = b ,
    \end{equation*}
    where $m > 0$, $c > 0$, $k > 0$, and $c^2 - 4km > 0$ (system is overdamped).
\end{exercise}

\begin{exercise}
    Using the Laplace transform solve
    \begin{equation*}
        m x'' + c x' + k x = 0 , \quad x(0) = a, \quad x'(0) = b ,
    \end{equation*}
    where $m > 0$, $c > 0$, $k > 0$, and $c^2 - 4km < 0$ (system is underdamped).
\end{exercise}

\begin{exercise}
    Using the Laplace transform solve
    \begin{equation*}
        m x'' + c x' + k x = 0 , \quad x(0) = a, \quad x'(0) = b ,
    \end{equation*}
    where $m > 0$, $c > 0$, $k > 0$, and $c^2 = 4km$ (system is critically damped).
\end{exercise}

\begin{exercise}
    Solve $x'' + x = u(t-1)$ for initial conditions $x(0) = 0$ and $x'(0) = 0$.
\end{exercise}

\begin{exercise}%
Solve $x''-x = (t^2-1) u(t-1)$ for initial conditions $x(0)=1$, $x'(0) = 2$ using the Laplace transform.
\end{exercise}
%\exsol{%
%$x(t) = (2e^{t-1}-t^2-1) u(t-1) -\frac{1}{2}e^{-t}+\frac{3}{2}e^t$
%}
 

\begin{exercise}
    Show the differentiation of the transform property.  Suppose $\mathcal{L} \bigl\{ f(t) \bigr\} = F(s)$, then show
    \begin{equation*}
        \mathcal{L} \bigl\{ -t f(t) \bigr\} = F'(s) .
    \end{equation*}
    Hint: Differentiate under the integral sign.
\end{exercise}

\begin{exercise}
    Solve $x''' + x = t^3 u(t-1)$ for initial conditions $x(0) = 1$ and $x'(0) = 0$, $x''(0) = 0$.
\end{exercise}

\begin{exercise}
    Show the second shifting property: $\mathcal{L} \bigl\{ f(t-a) \, u(t-a) \bigr\} = e^{-as} \mathcal{L} \bigl\{ f(t) \bigr\}$.
\end{exercise}

\begin{exercise}
    Let us think of the mass-spring system with a rocket from \exampleref{lt:rocketex}.  We noticed that the solution kept oscillating after the rocket stopped running.  The amplitude of the oscillation depends on the time that the rocket was fired (for 4 seconds in the example). 
    \begin{itemize}
    \item Find a formula for the amplitude of the resulting oscillation in terms of the amount of time the rocket is fired.
    \item Is there a nonzero time (if so what is it?) for which the rocket fires and the resulting oscillation has amplitude 0 (the mass is not moving)?
    \end{itemize}
\end{exercise}

\begin{exercise}
    Define
    \begin{equation*}
        f(t) =
        \begin{cases}
            {(t-1)}^2 & \text{if } \; 1 \leq t < 2, \\
            3-t & \text{if } \; 2 \leq t < 3, \\
            0 & \text{otherwise} .
        \end{cases}
    \end{equation*}
    \begin{itemize}
        \item Sketch the graph of $f(t)$.
        \item Write down $f(t)$ using the Heaviside function.
        \item Solve $x''+x=f(t)$, $x(0)=0$, $x'(0) = 0$ using Laplace transform.
    \end{itemize}
\end{exercise}

\begin{exercise}
    Find the transfer function for $m x'' + c x' + kx = f(t)$ (assuming the initial conditions are zero).
\end{exercise}

\begin{exercise}%
    Find the transfer function for $x' + x = f(t)$ (assuming the initial conditions are zero).
\end{exercise}
%\exsol{%
%$H(s) = \frac{1}{s+1}$
%}

%\setcounter{exercise}{100}


\end{document}